Edge effects and optimal field sizes in Alberta

### 2021-06-02 Manuscript/manuscript.Rmd
Worked on manuscript

- Field size and field boundaries are directly related to the conservation of SNL in agroecosystems
  - Size examples: wheat study from the UK
  - Boundary examples: flower strip studies, hedgerows
  - In North America, large financial incentives to make fields large and homogeneous, especially with large harvesting and planting equipment
  - Crop edge effects cause low yields at the margins of fields because of late emergence, poor microclimate, and competition with weeds
  - Ecosystem services should decay with distance from edge, so crops at the centre of a large field will not benefit from ecosystem services
  - Therefore, there should be a ``goldilocks" field size, where negative edge effects are canceled out by ecosystem services

@@

- In this paper we ask:
  1. How does crop yield mean and variance change with distance from the edge of field?
  2. Does this depend on type of field edge?
  3. Given this relationship, are there upper and lower bounds for field size if a farmer wanted to maximize average yield, or minimize variance?

@@

- Yield data was collected in discrete rectangles of the same length as the data interval (distance = speed $\times$ interval, typically 1 second) and the same width as the combine header (5-7 m)
  - We extracted the size of each polygon (m$^2$), dry yield (tonnes), and the spatial location, and the sequence of collection (1 - end of harvest)
  - Because of the large number of yield rectangles per field (30-800 thousand), we used the centroid of each polygon as its location
  - Seeding and application rates were constant with each field, so we did not consider inputs in our analysis

- Field boundaries were automatically digitized, then manually checked using satellite imagery from Google Earth and classified land cover data from AAFC 
  - Crop boundaries are flexible, and often change from year to year depending on planting and emergence conditions (e.g. flooding during some years)
  - Additionally, seminatural features often change from year to year
    - Ephemeral wetlands are flooded during some years, but consist mainly of grasses during dry years
    - Grass boundaries can change if fields are used for as haying or pasture during crop rotation
  - This makes accurate and consistent classification of field boundaries very difficult
  - Because of this, we define the following general categories for field boundaries:
    1. Standard: grassy field edge, staging yard, or road right-of-way (grassy strip typically 5-10m wide) 
    2. Wetland: permanent wetland; borders are largely unchanged from year-to-year
    3. Shelterbelt: permanent windbreaks, shelterbelts, remnant forests, or shrublands
    4. Other crop: annual crop or pasture with little or no visible boundary between planted areas
    5. Bare: unplanted, fallow, flooded area, temporary wetland, staging yard, oil and gas equipment, or road with no planted boundary
    6. Grassland: permanent seminatural grassland or pasture (not in rotation)

@@

- At each field, we fit an additive model of the effect of boundary distance on crop yield while accounting for within-field spatial variation and temporal variation in the combine yield monitor
  - Crop yield varies within a field due to soil conditions, moisture, seeding rates, herbicide application, and previous agricultural practices such as strip farming
  - While sensor calibration can reduce combine-level bias (such as a combine recording consistently higher/lower yields), this does not address sensor drift over time that occurs within fields
  - The yield monitors may record lower yields as sensors accumulate debris during harvest (pers. comm. Trent Clark), leading to changes in accuracy and bias over time
  - Additionally, ground speed is known to be extremely important to yield monitor accuracy [@arslan2002] 
  - To address this, we fit the following model:

  \begin{equation}
  log(yield) ~ Normal (\mu, \sigma)
  \mu = Intercept + log(Polygon Size) + f(Distance from Edge) + f(Easting, Northing) + f(Harvest Sequence)
  log(\sigma) = Intercept + log(Polygon Size) + f(Distance from Edge) + f(Easting, Northing) + f(Harvest Sequence)
  \end{equation}

>>>>

- Field size and field boundaries are directly related to the conservation of SNL in agroecosystems
  - Size examples: wheat study from the UK
  - Boundary examples: flower strip studies, hedgerows
  - In North America, there are financial incentives to make fields large and homogeneous, especially with large agricultural equipment
  - Crop edge effects cause low yields at the margins of fields because of late emergence, poor microclimate, and competition with weeds
  - At the same time, ecosystem services should decay with distance from edge, so crops at the centre of a large field will not benefit from ecosystem services
  - This suggests that yield may be maximized at distances between the edge and centre of the field, where ecosystem services cancel out negative edge effects

  <!-- - This suggests a ``goldilocks" field size, where negative edge effects are canceled out by ecosystem services -->

@@

- In this paper we ask:
  1. How does crop yield change with distance from the edge of field?
  2. Does this depend on type of field edge? 
  3. Is there an intermediate distance where yield is maximized or variance is minimized?

@@

- Yield data was collected in rectangles of the same length as the data interval (distance = speed $\times$ interval, typically 1 second) and the same width as the combine header (5-7 m)
  - We extracted the size of each polygon (m$^2$), dry yield (tonnes), and the spatial location, and the sequence of collection (1 - end of harvest)
  - Because of the large number of yield rectangles per field (30-800 thousand), we used the centroid of each polygon as its location, treating areal data as point data
  - Seeding and application rates were constant across fields, so we did not consider inputs in our analysis
  - We used dry yield (tonnes of seed/hectare after accounting for crop moisture) as our measure of crop yield
  - Due to the high number of samples at each field, we sub-sampled to 50,000 data points per field to reduce computation time

- Field boundaries were automatically digitized using buffers from the yield data locations, then manually checked using satellite imagery from Google Earth and classified land cover data from AAFC 
  - Crop boundaries are flexible, and often change yearly depending on planting and emergence conditions (e.g. flooding during some years)
  - Additionally, seminatural features often change yearly
    - Ephemeral wetlands are flooded during some years, but consist mainly of grasses during dry years
    - Grass boundaries can change if fields are used for as haying or pasture during crop rotation
  - This makes accurate and consistent classification of field boundaries difficult
  - We defined the following general categories for field boundaries:
    1. Standard: grassy field edge, staging yard, or road right-of-way (grassy strip typically 5-10m wide) 
    2. Wetland: permanent wetland; borders are largely unchanged from year-to-year
    3. Shelterbelt: permanent windbreaks, shelterbelts, remnant forests, or shrublands
    4. Other crop: annual crop or pasture with little or no visible boundary between planted areas
    5. Bare: unplanted, fallow, flooded area, temporary wetland (only present for a single season), staging yard, oil and gas equipment, or road without a planted boundary
    6. Grassland: permanent seminatural grassland or pasture (not in rotation)

@@

- At each field, we fit an additive model of the effect of boundary distance on crop yield while accounting for within-field spatial variation and temporal variation in the combine yield monitor
  - Ground speed is known to be extremely important to yield monitor accuracy [@arslan2002], with low ground speed registering higher yields 
  - Crop yield can vary within a field due to soil conditions, moisture, seeding rates, herbicide application, and previous agricultural practices 
    <!-- - such as strip farming -->
  - While sensor calibration can reduce combine-level bias (such as a combine recording consistently higher/lower yields across fields), this does not address sensor drift that occurs over time  within fields
  - This may be caused by sensors accumulating debris during harvest (pers. comm. Trent Clark), leading to changes in accuracy and bias over time

  - To model this, we fit the following model:

  \begin{equation}
  sqrt(yield) ~ Normal (\mu, \sigma)
  \mu = Intercept + log(Polygon Size) + f(Distance from Edge_i, b=12)_i + f(Easting, Northing, b=60) + f(Harvest Sequence, b=60)
  log(\sigma) = Intercept + log(Polygon Size) + f(Distance from Edge, b=12) + f(Easting, Northing, b=60) + f(Harvest Sequence, b=60)
  \end{equation}

  - where:
  1. Polygon Size = distance traveled x width of header bar ($m^2$)
  2. Distance from Edge = distance from field edge type _i_ (m)
  3. Easting, Northing = distance from centroid of field (m)
  4. Harvest Sequence = order of harvest within field (1--N points)
  5. f(x,b) = penalized thin-plate regression spline, where x is the predictor and b is the number of basis dimensions

- This model accounts for a) combine speed differences, b) edge effects, c) within-field spatial variation not related to edges, and d) temporal differences in harvest 
  - Spatial or temporal variation is typically modelled using a Gaussian Process Model (e.g. Kriging), but this was computationally infeasible with 50,000 data points per field
  - Penalized regression splines offer a compromise, as they account for nonlinear ``wiggly" relationships in the same way as Gaussian processes but with substantially reduced computation time
  - The number of basis dimensions was checked with the _gam.check_ function from _mgcv_
  - A smoother was not used for polygon size because the relationship was easily approximated with a log-linear relationship using a single slope term
  - All models were fit in _R_ using the _mgcv_ library, and figures were created with _ggplot2_ and _ggpubr_

- Most models still had a large amount of spatio-temporal autocorrelation in residuals
  - We basically just ignored this

- To combine results from all field-level models while also accounting for field-level uncertainty, we used an approach similar to bootstrapping of hierarchical mixed effects models
  - The approach was as follows
  1. Extract posterior samples of smoother parameters from each field-level model
  2. Fit meta-model of all smoothers, and save parameters
  3. Repeat 1000 times, and calculate credible intervals (10-90th percentiles of draws)

- We compared this
